---
title: "ECON 424 final project R"
author: "Kye Jones"
date: "2023-12-06"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message = FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)
library(fastDummies)
library(wooldridge)
library(lmtest)
library(sandwich)
library(ggplot2)
library(tidyverse)
library(car)
library(ggfortify)
library(mice)
```

```{r}
library(readxl)
ECON_424_Paper_Data <- read_excel("ECON 424 Paper Data.xlsx")
View(ECON_424_Paper_Data)
```


```{r}
summary(ECON_424_Paper_Data)
head(ECON_424_Paper_Data)
str(ECON_424_Paper_Data)
```

ESG Overall: The scores range from 1.002 to 3.454 with a median of 1.608, indicating some firms have very high ESG performance. There's a significant spread in the data, as seen by the range.
ESG Social: These scores are slightly higher on average than overall ESG scores, with a median of 1.725 and a similar broad range, suggesting a focus on social aspects.
ESG Environmental: The environmental scores are lower on average compared to social scores, with a median of 1.419, indicating a potential area for improvement for many firms.
Industry: The two-digit industry codes have some missing values, showing a diverse range of industries represented.
Female Proportion in Industry (indus_femprop): Averages at around 38% but ranges from 11% to 54.9%, indicating variability in female employment across industries.
Dirty Industry (dirtyindus): Around 27% of the firms are classified in 'dirty' industries, potentially affecting their ESG scores.
Firm Leverage: On average, firms have a leverage ratio of about 27.8%, but the range is wide, suggesting varying financial risk profiles.
Capital Intensity: The average is low at 7.5%, but there's a high maximum value, indicating some firms are highly capital-intensive.
Firm Size: Represented by the log of total assets, showing firms range from small to very large in size.
Board Size: The board size varies significantly, suggesting different governance structures.
Number of Executives: Some firms have up to four executives on their board, indicating diverse management structures.
Gender Diversity: There's a wide variation in board gender diversity, with a mean of 19.28%, suggesting many boards are male-dominated.
Return on Assets (ROA): Varies widely with some firms having negative ROA, indicating loss-making operations for some of the sampled period.
Age Diversity: There's a significant range in the age diversity of boards, with some having a very wide range of ages represented.
Board Independence: Many firms have fully independent boards, with a mean independence level of 83.24%.
Missing Data: There are missing values in several variables, including industry classification, industry female proportion, firm leverage, and cap intensity. The missing data should be addressed before running your regression analysis.


```{r}
colSums(is.na(ECON_424_Paper_Data))
```

```{r}
cleaned_data <- ECON_424_Paper_Data

cleaned_data <- cleaned_data[!is.na(cleaned_data$indus_twodigit), ]


vars_to_impute <- c("indus_femprop", "dirtyindus", "firmleverage", "capintensity", "firmsize", "roa", "agediversity")
imputed_data <- mice(cleaned_data[, vars_to_impute], method = 'median', m = 1, maxit = 1)
cleaned_data[, vars_to_impute] <- complete(imputed_data)

sapply(cleaned_data, function(x) sum(is.na(x)))
```

This code excludes observations with missing industry codes, imputes the median for continuous financial variables and age diversity, and then checks to confirm there are no remaining missing values. 




```{r}
# Model 1: ESG scores regressed on gender diversity and control variables
model1 <- lm(esg_overall ~ gendiversity + indus_femprop + firmsize + firmleverage + agediversity + boardindepend + capintensity + boardsize + numexecutive + indus_twodigit + dirtyindus, data = ECON_424_Paper_Data)
summary(model1)

# Model 2: Financial performance (ROA) regressed on gender diversity and control variables
model2 <- lm(roa ~ gendiversity + indus_femprop + firmsize + firmleverage + agediversity + boardindepend + capintensity + boardsize + numexecutive + indus_twodigit + dirtyindus, data = ECON_424_Paper_Data)
summary(model2)

```

```{r}
plot(model1)
plot(residuals(model1) ~ fitted(model1))
vif(model1)
```

```{r}
plot(model2)
plot(residuals(model2) ~ fitted(model2))
vif(model2)
```

```{r}
model3 <- lm(esg_overall ~ gendiversity * firmsize + indus_femprop + firmleverage + agediversity + boardindepend + capintensity + boardsize + numexecutive + indus_twodigit + dirtyindus, data = ECON_424_Paper_Data)
summary(model3)
```

```{r}
interaction.plot(ECON_424_Paper_Data$firmsize, ECON_424_Paper_Data$gendiversity, ECON_424_Paper_Data$esg_overall)
```

```{r}
plot(residuals(model3) ~ fitted(model3))
qqnorm(residuals(model3))
qqline(residuals(model3))
vif(model3)
plot(model3)
```

```{r}
summary_model1 <- coeftest(model1, vcov = vcovHC(model1, type = "HC1"))
print(summary_model1)
summary_model2 <- coeftest(model2, vcov = vcovHC(model2, type = "HC1"))
print(summary_model2)
summary_model3 <- coeftest(model3, vcov = vcovHC(model3, type = "HC1"))
print(summary_model3)
```

```{r}
# Joint significance tests for models
# Model 1: Test if gender diversity is significant
linearHypothesis(model1, "gendiversity = 0")

# Model 2: Test if gender diversity is significant
linearHypothesis(model2, "gendiversity = 0")

# Model 3: Test if both gender diversity and the interaction with firm size are significant
linearHypothesis(model3, c("gendiversity = 0", "firmsize = 0", "gendiversity:firmsize = 0"))
```

```{r}
# Model comparison using AIC and BIC
model_comparison <- data.frame(
  Model = c("Model 1", "Model 2", "Model 3"),
  AIC = c(AIC(model1), AIC(model2), AIC(model3)),
  BIC = c(BIC(model1), BIC(model2), BIC(model3))
)
print(model_comparison)
```

```{r}
predicted_values1 <- predict(model1, newdata = ECON_424_Paper_Data)


length_actual <- length(ECON_424_Paper_Data$esg_overall)
length_predicted <- length(predicted_values1)


if (length_actual == length_predicted) {
  plot(ECON_424_Paper_Data$esg_overall, predicted_values1, 
       main = "Predicted vs Actual Values",
       xlab = "Actual ESG Scores",
       ylab = "Predicted ESG Scores",
       pch = 16)
  abline(0, 1, col = "red", lwd = 2) # Add a red line at 45 degrees
} else {
  stop("The lengths of actual and predicted values do not match.")
}
```

```{r}
predicted_values2 <- predict(model2, newdata = ECON_424_Paper_Data)
if (length(ECON_424_Paper_Data$roa) == length(predicted_values2)) {
  plot(ECON_424_Paper_Data$roa, predicted_values2, 
       main = "Predicted vs Actual Values",
       xlab = "Actual ROA",
       ylab = "Predicted ROA",
       pch = 16)
  abline(0, 1, col = "blue", lwd = 2) # Add a blue line at 45 degrees
} else {
  stop("The lengths of actual and predicted values do not match.")
}

```

```{r}

autoplot(model1, which = 1:2, ncol = 2)
autoplot(model2, which = 1:2, ncol = 2)
autoplot(model3, which = 1:2, ncol = 2)
```

```{r}

coeftest_robust_model1 <- coeftest(model1, vcov = vcovHC(model1, type = "HC1"))
coeftest_robust_model2 <- coeftest(model2, vcov = vcovHC(model2, type = "HC1"))
coeftest_robust_model3 <- coeftest(model3, vcov = vcovHC(model3, type = "HC1"))
```


```{r}

print("Robust Standard Errors for Model 1")
print(coeftest_robust_model1)

print("Robust Standard Errors for Model 2")
print(coeftest_robust_model2)

print("Robust Standard Errors for Model 3")
print(coeftest_robust_model3)

```

